Higgs Assisted Razor Search for Higgsinos at 100 TeV

### 2019-07-23 main.tex
Updates from Overleaf

\documentclass[a4paper,11pt]{article}
 \providecommand{\tightlist}{%
     \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{graphicx}
\usepackage{jheppub} 
\usepackage{comment} 
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{slashed}
\usepackage{pgfplots}
\usepackage{booktabs, longtable}
\bibliographystyle{JHEP.bst}

\newcommand{\Shufang}[1]{{\bf\color{Maroon}  #1}}
\newcommand{\Adarsh}[1]{{\bf\color{RoyalBlue} AP: #1}}
\newcommand{\ap}[1]{\textcolor{RoyalBlue}{#1}}
\renewcommand{\H}{\widetilde{H}^0}
\newcommand{\B}{\widetilde{B}^0}
\newcommand{\N}{\widetilde{\chi}^0}
\newcommand{\C}{\widetilde{\chi}^\pm}
\newcommand{\image}[2]{\parbox{#1\textwidth}{\includegraphics[width=#1\textwidth]{images/#2.pdf}}}
\newcommand{\met}{{\slashed{E}_T}}

@@

Since the Bino is the only supersymmetric particle lighter than the Higgsino in
the mass hierarchy we consider, the neutral Higgsino will decay to a Bino and a
neutral SM boson (\emph{Z/h}) with a branching ratio of 100\%.  Searches for
Higgsino NLPs with a Bino LSP  via  multi-lepton final states have been studied
in Ref.~\cite{Gori:2014oua} at a 100 TeV $pp$ collider, for pair production of
$\N_{2,3}\C_1$ and
$\widetilde\chi_{1}^+\widetilde\chi_1^-$. In this paper, we attempt to exploit
the relatively high branching ratio of the SM Higgs boson to \emph{b} quarks,
and study the potential of the channel of $\N_2\N_3\rightarrow Zh
\N_1\N_1\rightarrow bb\ell\ell \met$.

% Why a 100 TeV collider?
A 100 TeV $pp$ collider would open up an immense number of physics
opportunities not afforded to a 14 TeV LHC, including exploring the most
interesting regions of the split SUSY parameter space
\cite{Arkani-Hamed:2015vfh}.
% Status of 100 TeV planning
Such a collider would be a natural next step after the LHC, and is being
actively discussed in the particle physics community, with two major proposals
being the Future Circular Collider (FCC)-hh by CERN \cite{FCC-hh}, and the
Super Proton Proton Collider (SppC)  in China \cite{CEPC}.  As the next energy
frontier machine, it is of crucial importance to fully explore the physics
potential of a 100 TeV $pp$ collider, especially for new particles that are
either too heavy or too rare to be produced at the
LHC~\cite{Arkani-Hamed:2015vfh,Contino:2016spe,Golling:2016gvc,Mangano:2016jyj}.
In particular, studies have already been performed on the prospects of
discovering neutralino dark matter for compressed \cite{Low:2014cba,
diCortona:2014yua,Cirelli:2014dsa,Mahbubani:2018tin,Han:2018wus} and
well-separated \cite{Gori:2014oua,Acharya:2014pua} neutralino spectra.

@@

At a 100 TeV collider, the  SM backgrounds are going to be even larger than
at the LHC.  An ancillary goal of this paper is to investigate the
potential for machine learning (ML) techniques to augment our analysis.
With the advent of more powerful computers and simultaneous advances in
the field of statistical learning in recent years, the usage of ML is
rising in experimental particle physics -- in fact, the discovery of the
SM Higgs boson in 2012 was done with the help of neural networks
\cite{Aad:2012tfa} and boosted decision trees (BDT)
\cite{Chatrchyan:2012xdj}.

The rest of the paper is structured as follows. In Sec.~\ref{sec:model}, we
describe our model and search channel in more detail, and list the existing
experimental search bounds. In Sec.~\ref{sec:analysis}, we describe our
analysis strategies for both traditional cut-and-count analyses and analyses
performed using BDT. In Sec.~\ref{sec:results}, we present the results of our
analyses, i.e. the expected 5 $\sigma$ discovery and 95\% C.L. exclusion
reaches. Finally, we conclude with the implications in
Sec.~\ref{sec:conclusion}.

@@

The neutralino sector of MSSM consists of four mass eigenstates
($\N_1,\N_2,\N_3,\N_4$), which are mixtures of the Bino, Wino, and two neutral
Higgsinos: ($\B,\widetilde{W}^0,\H_d,\H_u$). In the basis of these gauge
eigenstates, the mass matrix of the neutralinos can be written as

\newcommand{\cb}{ c_\beta}
\newcommand{\cw}{ c_W}
\newcommand{\sinb}{ s_\beta}
\newcommand{\sw}{ s_W}
\newcommand{\mz}{ m_Z}

@@

\noindent where we employ the notation $s_W, c_W = \sin\theta_W, \cos\theta_W$
for $\theta_W$ being the Weinberg mixing angle. $s_\beta, c_\beta$ are defined
accordingly with $\tan\beta= v_u/v_d$, where $v_u$ and $v_d$ are the vacuum
expectation values of $H_u^0$ and $H_d^0$, with $v_u^2+v_d^2=v^2=(246\ {\rm
GeV})^2$. The parameters $M_1$, $M_2$, and $\mu$ are the mass parameters of
Bino, Wino, and Higgsinos.  The mass eigenstates are usually labeled as
$\chi_i^0$ for $i=1 \ldots 4$ with increasing mass eigenvalues.

In the limit of $\mz \ll |\mu\pm M_1|, |\mu\pm M_2|,$ the mass eigenstates
are, to good approximation, a nearly pure Bino, $\B$, with mass $M_1$,
a nearly pure Wino, $\widetilde{W}^0$, with mass $M_2$, and nearly pure Higgsinos
$\H_{1,2} = (\H_u \pm \H_d)/\sqrt{2}$, with mass
$|\mu|$.

@@

\noindent where $\widetilde{\chi}_{i,j}$ can be a neutralino or chargino, $\N_1$
is the lightest neutralino, and $V$ being any of the SM gauge bosons $W$, $Z$
or scalar $h$.  These searches work well since the large mass difference
between the electroweakinos can lead to energetic leptons that can be easily
identified. Among the multi-lepton searches for Higgsinos, the trilepton
searches: $\N_{2,3}\C_1\pm\rightarrow WZ\N_1\N_1\rightarrow \ell\ell\ell +
\met$, with $\ell=e, \ \mu$,  have the best reach, due to the relatively high
production cross-section of chargino-neutralino pairs combined with the large
reduction in $t\overline{t}$ and QCD backgrounds obtained by requiring three
leptons.  Higgsino NLSPs can be excluded at 95\% C.L. with mass up to 2.8 TeV
or be discovered with  mass up to 1.8 TeV, for a massless Bino LSP.   The reach
reduces for larger $M_1$, which vanishes for $M_1\gtrsim 600$ GeV (1000 GeV)
for discovery (exclusion).    Multilepton searches with \emph{Zh} and \emph{ZZ}
as the intermediate dibosons are unlikely to be as powerful, due to the smaller
pair-production cross-section of neutral Higgsinos, combined with the low
branching ratio of \emph{Z/h} to leptons. However, if we move beyond
multilepton searches, the channel with \emph{Zh} (originally suggested in
\cite{Han:2013kza}) as the intermediate dibosons emerges as a possible
competitor to the \emph{WZ} channel.

In the left panel of Fig.~\ref{fig:xsection_plot}, we show the pair-production
cross section $\N_2\N_3$ for the Higgsino-like NLSPs.  The cross section varies
between 200 fb to 0.5 fb for Higgsino masses between 500 GeV and 2500 GeV. The
right panel of Fig.~\ref{fig:xsection_plot} shows the decay branching fractions
of $\chi_{2,3}^0\rightarrow Z/h \chi_1^0$, as calculated in SUSY-HIT
\cite{Djouadi:2006bz}.  While one of the Higgsinos mostly decays to $Z$, the
other Higgsino mostly decays to $h$.    Therefore,  for pair produced Higgsinos
$\N_2\N_3$, the \emph{Zh} mode has the greatest branching fraction.  After
multiplying all the branching fractions, the signal cross sections of

@@

\noindent become comparable  to trilepton processes, as shown in table
\ref{tab:xsections}. $Zh+\met$ signal thus provide an alternative discovery
channel for Higgsinos. The signal contains two \emph{b}-jets, two same flavor,
opposite sign leptons, and missing transverse energy. The main backgrounds for
this process are: $t\overline{t}$, \emph{tbW} with the $b$ and $W$ not coming
from a \emph{t}, and \emph{bbWW} with no intermediate $s$-channel top quarks.

@@

\begin{table}
  \centering
  \begin{tabular}{l|rr}
    \toprule
    Stage & $\N_{2,3}\C_1$ & $\N_2\N_3$\\
    \midrule
    Pair production cross section & 60 fb & 17.2 fb\\
    Intermediate diboson contribution & ($WZ$) 29.7 fb  & ($Zh$) 9.15 fb \\
    Applying \emph{BR}$(W\rightarrow \ell\nu)$, \emph{BR}$(Z\rightarrow \ell\ell)$ \& \emph{BR}$(h\rightarrow bb)$ & 0.42 fb & 0.37 fb\\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Higgsino pair production cross-sections for
    $(\N_{2,3}\C_1)$ and
    $(\N_{2}\N_{3})$, for $\mu\approx 1$ TeV, at a 100
    TeV \emph{pp} collider. 
    The pair-production cross section for $\N_{2,3}\C_1$ is taken from
    \cite{Gori:2014oua}, while the pair-production cross-section for $\N_2\N_3$
    is calculated at Next-to-Leading Order (NLO) using Prospino2. The branching
    ratios to the diboson intermediate states are calculated using SUSY-HIT,
    and the branching ratios to the SM final states are taken from
    \cite{Olive:2016xmw}.}
  \label{tab:xsections}
\end{table}

Searches for pair produced nearly mass-degenerate Higgsinos  have been
performed at ATLAS and CMS for various decay  topologies
\cite{Aaboud:2018htj,Sirunyan:2018ubx,Aaboud:2018zeb}.  
For neutralino-chargino pair production with $WZ$ and $WH$ topologies, the 95\%
C.L. exclusion limits are about 650 GeV, 490 GeV, and 535 GeV for $WZ$, $WH$
and an equal mixture of $WZ+WH$ topology, with the most strongest limits from
the multi-lepton searches.  For neutralino pair production with $ZZ$, $ZH$ and
$HH$ topologies~\footnote{While such searches assume gravitino LSP with
Higgsino NLSP, the limits are still relevant for our Higgsino NLSP with Bino
LSP scenario.}, the limits are about 650 $-$ 750 GeV via multi-lepton or
multi-$b$ jets searches, depending on the decay branching fractions of Higgsino
NLSP.

>>>>

\documentclass[a4paper,11pt]{article}
 \providecommand{\tightlist}{%
     \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{graphicx}
\usepackage{jheppub} 
\usepackage{comment} 
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{slashed}
\usepackage{pgfplots}
\usepackage{booktabs, longtable}
\usepackage{microtype}
\bibliographystyle{JHEP.bst}

\newcommand{\Shufang}[1]{{\bf\color{Maroon}  #1}}
\newcommand{\Adarsh}[1]{{\bf\color{RoyalBlue} AP: #1}}
\newcommand{\ap}[1]{\textcolor{RoyalBlue}{#1}}
\renewcommand{\H}{\widetilde{H}^0}
\newcommand{\W}{\widetilde{W}^0}
\newcommand{\B}{\widetilde{B}^0}
\newcommand{\N}{\widetilde{\chi}^0}
\newcommand{\C}{\widetilde{\chi}^\pm}
\newcommand{\image}[2]{\parbox{#1\textwidth}{%
\includegraphics[width=#1\textwidth]{images/#2.pdf}}
}
\newcommand{\met}{{\slashed{E}_T}}

@@

Since the Bino is the only supersymmetric particle lighter than the Higgsino in
the mass hierarchy we consider, the neutral Higgsino will decay to a Bino and a
neutral SM boson (\emph{Z/h}) with a branching fraction of 100\%.  Searches for
Higgsino NLPs with a Bino LSP  via  multi-lepton final states have been studied
in Ref.~\cite{Gori:2014oua} at a 100 TeV $pp$ collider, for pair production of
$\N_{2,3}\C_1$ and
$\widetilde\chi_{1}^+\widetilde\chi_1^-$. In this paper, we attempt to exploit
the relatively high branching fraction of the SM Higgs boson to \emph{b} quarks,
and study the potential of the $\N_2\N_3\rightarrow Zh
\N_1\N_1\rightarrow bb\ell\ell \met$ channel.

% Why a 100 TeV collider?
A 100 TeV $pp$ collider would open up an immense number of physics
opportunities not afforded to the 14 TeV LHC, including exploring the most
interesting regions of the split SUSY parameter space
\cite{Arkani-Hamed:2015vfh}.
% Status of 100 TeV planning
Such a collider would be a natural next step after the LHC, and is being
actively discussed in the particle physics community, with two major proposals
being the Future Circular Collider (FCC)-hh by CERN \cite{FCC-hh}, and the
Super Proton Proton Collider (SppC)  in China \cite{CEPC}.  As the next energy
frontier machine, it is crucial to fully explore the physics
potential of a 100 TeV $pp$ collider, especially for new particles that are
either too heavy or too rare to be produced at the
LHC~\cite{Arkani-Hamed:2015vfh,Contino:2016spe,Golling:2016gvc,Mangano:2016jyj}.
In particular, studies have already been performed on the prospects of
discovering neutralino dark matter for compressed \cite{Low:2014cba,
diCortona:2014yua,Cirelli:2014dsa,Mahbubani:2018tin,Han:2018wus} and
well-separated \cite{Gori:2014oua,Acharya:2014pua} neutralino spectra.

@@

At a 100 TeV collider, the  SM backgrounds are going to be even larger than
at the LHC.  An ancillary goal of this paper is to investigate the
potential for machine learning (ML) techniques to augment our analysis.
With the advent of more powerful computers and simultaneous advances in
the field of statistical learning in recent years, the usage of ML is
rising in experimental particle physics -- in fact, the discovery of the
SM Higgs boson in 2012 was done with the help of neural networks
\cite{Aad:2012tfa} and boosted decision trees (BDTs)
\cite{Chatrchyan:2012xdj}.

The rest of the paper is structured as follows. In Sec.~\ref{sec:model}, we
describe our model and search channel in more detail, and list the existing
experimental search bounds. In Sec.~\ref{sec:analysis}, we describe our
analysis strategies for both traditional cut-and-count analyses and analyses
performed using BDTs. In Sec.~\ref{sec:results}, we present the results of our
analyses, i.e. the expected 5 $\sigma$ discovery and 95\% C.L. exclusion
reaches. Finally, we conclude with the implications in
Sec.~\ref{sec:conclusion}.

@@

The neutralino sector of MSSM consists of four mass eigenstates
($\N_1,\N_2,\N_3,\N_4$), which are mixtures of the Bino, Wino, and two neutral
Higgsinos: ($\B,\W,\H_d,\H_u$). In the basis of these gauge
eigenstates, the mass matrix of the neutralinos can be written as

\newcommand{\cb}{ c_\beta}
\newcommand{\cw}{ c_{\theta_W}}
\newcommand{\sinb}{ s_\beta}
\newcommand{\sw}{ s_W}
\newcommand{\mz}{ m_Z}

@@

\noindent where we employ the notation $s_\theta, c_\theta = \sin\theta, \cos\theta$. Here,  $\theta_W$ is the Weinberg mixing angle, 
and $\tan\beta= v_u/v_d$, where $v_u$ and $v_d$ are the vacuum
expectation values of $H_u^0$ and $H_d^0$, with $v_u^2+v_d^2=v^2=(246\ {\rm
GeV})^2$. The parameters $M_1$, $M_2$, and $\mu$ are the mass parameters of
the Bino, Wino, and the Higgsino, respectively.
The mass eigenstates are labeled as $\N_i$ for $i=1 \ldots 4$ with increasing mass eigenvalues.
In the limit where $\mz \ll |\mu\pm M_1|, |\mu\pm M_2|,$ the mass eigenstates
are, to good approximation, a nearly pure Bino, $\B$, with mass $M_1$,
a nearly pure Wino, $\W$, with mass $M_2$, and nearly pure Higgsinos
$\H_{1,2} = (\H_u \pm \H_d)/\sqrt{2}$, with mass
$|\mu|$.

@@

\noindent where $\widetilde{\chi}_{i,j}$ can be a neutralino or chargino, $\N_1$
is the lightest neutralino, and $V$ being any of the SM gauge bosons $W$, $Z$
or scalar $h$.  These searches work well since the large mass difference
between the electroweakinos can lead to energetic leptons that can be easily
identified. Among the multi-lepton searches for Higgsinos, the trilepton
searches: $\N_{2,3}\C_1\rightarrow WZ\N_1\N_1\rightarrow \ell\ell\ell +
\met$, with $\ell=e, \ \mu$,  have the best reach, due to the relatively high
production cross-section of chargino-neutralino pairs combined with the large
reduction in $t\overline{t}$ and QCD backgrounds from requiring three
leptons.  Higgsino NLSPs can be excluded at 95\% C.L. with mass up to 2.8 TeV
or be discovered with  mass up to 1.8 TeV, for a massless Wino LSP.   The reach
reduces for larger $M_1$, and vanishes for $M_1\gtrsim 600$ GeV (1000 GeV)
for discovery (exclusion).  Multilepton searches with \emph{Zh} and \emph{ZZ}
as the intermediate dibosons are unlikely to be as powerful, due to the smaller
pair-production cross-section of neutral Higgsinos, combined with the low
branching fraction of \emph{Z/h} to leptons. However, if we move beyond
multilepton searches, the channel with \emph{Zh} (originally suggested in
Ref.~\cite{Han:2013kza}) as the intermediate dibosons emerges as a possible
competitor to the \emph{WZ} channel.

In the left panel of Fig.~\ref{fig:xsection_plot}, we show the pair-production
cross section $\N_2\N_3$ for the Higgsino-like NLSPs.  The cross section varies
between 200 fb to 0.5 fb for Higgsino masses between 500 GeV and 2500 GeV. The
right panel of Fig.~\ref{fig:xsection_plot} shows the decay branching fractions
of $\chi_{2,3}^0\rightarrow Z/h \chi_1^0$, as calculated by SUSY-HIT
\cite{Djouadi:2006bz}.  While one of the Higgsinos mostly decays to $Z$, the
other Higgsino mostly decays to $h$.    Therefore,  for pair produced Higgsinos
$\N_2\N_3$, the \emph{Zh} mode has the greatest branching fraction.  After
multiplying all the branching fractions, the signal cross sections of

@@

\noindent become comparable  to trilepton processes, as shown in Table
\ref{tab:xsections}. Thus, the $Zh+\met$ signal provides an alternative discovery
channel for Higgsinos. The signal contains two \emph{b}-jets, two same flavor,
opposite sign leptons, and missing transverse energy. The main backgrounds for
this process are: $t\overline{t}$, \emph{tbW} with the $b$ and $W$ not coming
from a \emph{t}, and \emph{bbWW} with no intermediate $s$-channel top quarks.

@@

\begin{table}
  \centering
  \begin{tabular}{l|rr}
    \toprule
    Stage & $\N_{2,3}\C_1$ & $\N_2\N_3$\\
    \midrule
    Pair production cross section & 60 fb & 17.2 fb\\
    Intermediate diboson contribution & ($WZ$) 29.7 fb  & ($Zh$) 9.15 fb \\
    Applying \emph{BR}$(W\rightarrow \ell\nu)$, \emph{BR}$(Z\rightarrow \ell\ell)$ \& \emph{BR}$(h\rightarrow bb)$ & 0.42 fb & 0.37 fb\\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Higgsino pair production cross-sections for
    $(\N_{2,3}\C_1)$ and
    $(\N_{2}\N_{3})$, for $\mu\approx 1$ TeV, at a 100
    TeV \emph{pp} collider. 
    The pair-production cross section for $\N_{2,3}\C_1$ is taken from
    \cite{Gori:2014oua}, while the pair-production cross-section for $\N_2\N_3$
    is calculated at Next-to-Leading Order (NLO) using Prospino2. The branching
    ratios to the diboson intermediate states are calculated using SUSY-HIT,
    and the branching fractions to the SM final states are taken from
    \cite{Olive:2016xmw}.}
  \label{tab:xsections}
\end{table}

Searches for pair produced nearly mass-degenerate Higgsinos  have been
performed at ATLAS and CMS for various decay  topologies
\cite{Aaboud:2018htj,Sirunyan:2018ubx,Aaboud:2018zeb}.  
For neutralino-chargino pair production with \emph{WZ} and \emph{WH} topologies, the 95\%
C.L. exclusion limits are about 650 GeV, 490 GeV, and 535 GeV for \emph{WZ}, \emph{WH}
and an equal mixture of \emph{WZ}+\emph{WH} topology, with the strongest limits coming from
the multi-lepton searches.  For neutralino pair production with \emph{ZZ, ZH}, and \emph{HH} topologies~\footnote{While such searches assume gravitino LSP with
Higgsino NLSP, the limits are still relevant for our Higgsino NLSP with Bino
LSP scenario.}, the limits are about 650 $-$ 750 GeV via multi-lepton or
multi-$b$ jets searches, depending on the decay branching fractions of Higgsino
NLSP.
